Different forms of uncertainty differently affect the evolution of social learning

### 2022-08-29 meanNetPayoffs.tex
updated legend in payoff fig
major figure updates

\usepackage{geometry}
\geometry{
paperwidth=8in,
paperheight=6.25in,
margin=0.25in
}

@@

\begin{document}

    \begin{minipage}{3.75in}
      \centering
      {\hspace{5.25em}\huge $B = 2$}
    \end{minipage}%
    \begin{minipage}{3.75in}
      \centering
      {\hspace{2.0em}\huge $B = 10$}
    \end{minipage}~\\

    \begin{minipage}{3.75in}
    \begin{rotate}{90}
      {\parbox{2.5in}{
          \centering
          \vspace{-2.5em} {\huge$ \pilow = 0.1$} \\
          {\begin{rotate}{-90}{\huge $\frac{\meanpi}{L}$}\hspace{3em}\end{rotate}}
      }}
    \end{rotate}%
    \hspace{2em}
      \includegraphics[width=\textwidth]{Figures/mean_prev_net_payoff_over_u_lowpayoff=0.1_nbehaviors=2.pdf}
    \end{minipage}\noindent\begin{minipage}{3.75in}%
      \includegraphics[width=\textwidth]{Figures/mean_prev_net_payoff_over_u_lowpayoff=0.1_nbehaviors=10.pdf}
    \end{minipage}~\\[0.5em]

    \begin{minipage}{3.75in}
    \begin{rotate}{90}
      {\parbox{2.5in}{
          \centering
          \vspace{-2.5em} {\huge$ \pilow = 0.45$} \\
          {\begin{rotate}{-90}{\huge $\frac{\meanpi}{L}$}\hspace{3em}\end{rotate}}
      }}
    \end{rotate}%
    \hspace{2em}
      \includegraphics[width=\textwidth]{Figures/mean_prev_net_payoff_over_u_lowpayoff=0.45_nbehaviors=2.pdf}
        \\[-2.75em]
        \begin{center}
          {\hspace{3.25em} \huge $\quad u$}
      \end{center}
        \end{minipage}\noindent\begin{minipage}{3.75in}%
      \includegraphics[width=\textwidth]{Figures/mean_prev_net_payoff_over_u_lowpayoff=0.45_nbehaviors=10.pdf}
      \\[-2.75em]
      \begin{center}
        {\huge $\quad u$}
      \end{center}
    \end{minipage}
\end{document}

>>>>

\usepackage{geometry}
\geometry{
paperwidth=12in,
paperheight=12in,
margin=0.25in
}

@@

\begin{document}
\begin{tikzpicture}
      \draw[->,thick] (-.1,0)--(1.15,0) node[above]
        {\huge $B$};
        % {Number of behaviors $B$};
      \draw[->,thick] (0,.1)--(0,-1.15) node[below]
        {\huge $\pilow$};
        % {Low payoff frequency $\pilow$};
	  \end{tikzpicture}~\\[-2em]

    \begin{minipage}{3.75in}
      \centering
      {\hspace{5.25em}\huge $B = 2$}
    \end{minipage}%
    \begin{minipage}{3.75in}
      \centering
      {\hspace{1.0em}\huge $B = 4$}
    \end{minipage}%
    \begin{minipage}{3.75in}
      \centering
      {\hspace{2.0em}\huge $B = 10$}
    \end{minipage}~\\

    \begin{minipage}{3.75in}
    \begin{rotate}{90}
      {\parbox{2.5in}{
          \centering
          \vspace{-2.5em} {\huge$ \pilow = 0.1$} \\
          {\begin{rotate}{-90}{\huge $\frac{\meanpi}{L}$}\hspace{3em}\end{rotate}}
      }}
    \end{rotate}%
    \hspace{2em}
      \includegraphics[width=\textwidth]{Figures/mean_prev_net_payoff_over_u_lowpayoff=0.1_nbehaviors=2.pdf}
    \end{minipage}\noindent\hspace{1.25em}
	\begin{minipage}{3.75in}%
      \includegraphics[width=\textwidth]{Figures/mean_prev_net_payoff_over_u_lowpayoff=0.1_nbehaviors=4.pdf}
    \end{minipage}\noindent
	\begin{minipage}{3.75in}%
      \includegraphics[width=\textwidth]{Figures/mean_prev_net_payoff_over_u_lowpayoff=0.1_nbehaviors=10.pdf}
    \end{minipage}~\\[0.5em]

    \begin{minipage}{3.75in}
    \begin{rotate}{90}
      {\parbox{2.5in}{
          \centering
          \vspace{-2.5em} {\huge$ \pilow = 0.45$} \\
          {\begin{rotate}{-90}{\huge $\frac{\meanpi}{L}$}\hspace{3em}\end{rotate}}
      }}
    \end{rotate}%
    \hspace{2em}
      \includegraphics[width=\textwidth]{Figures/mean_prev_net_payoff_over_u_lowpayoff=0.45_nbehaviors=2.pdf}
	\end{minipage}\noindent\hspace{1.25em}
	\begin{minipage}{3.75in}%
      \includegraphics[width=\textwidth]{Figures/mean_prev_net_payoff_over_u_lowpayoff=0.45_nbehaviors=4.pdf}
    \end{minipage}\noindent
	\begin{minipage}{3.75in}%
      \includegraphics[width=\textwidth]{Figures/mean_prev_net_payoff_over_u_lowpayoff=0.45_nbehaviors=10.pdf}
    \end{minipage}~\\[0.5em]

    \begin{minipage}{3.75in}
    \begin{rotate}{90}
      {\parbox{2.5in}{
          \centering
          \vspace{-2.5em} {\huge$ \pilow = 0.8$} \\
          {\begin{rotate}{-90}{\huge $\frac{\meanpi}{L}$}\hspace{3em}\end{rotate}}
      }}
    \end{rotate}%
    \hspace{2em}
      \includegraphics[width=\textwidth]{Figures/mean_prev_net_payoff_over_u_lowpayoff=0.8_nbehaviors=2.pdf}
        \\[-2.75em]
        \begin{center}
          {\hspace{3.25em} \huge $\quad u$}
      \end{center}
	  \end{minipage}\noindent\hspace{1.25em}
		\begin{minipage}{3.75in}%
		  \includegraphics[width=\textwidth]{Figures/mean_prev_net_payoff_over_u_lowpayoff=0.8_nbehaviors=4.pdf}
		  \\[-2.75em]
	  \begin{center}
        {\huge $\quad u$}
      \end{center}
    \end{minipage}
	\begin{minipage}{3.75in}%
		  \includegraphics[width=\textwidth]{Figures/mean_prev_net_payoff_over_u_lowpayoff=0.8_nbehaviors=10.pdf}
		  \\[-2.75em]
	  \begin{center}
        {\huge $\quad u$}
      \end{center}
    \end{minipage} \\[2.5em]

    \begin{center}
      \includegraphics[width=1.25in]{Figures/legendElements/solid.pdf}
      {\Huge Simulations} \quad \quad 
      \includegraphics[width=1.25in]{Figures/legendElements/ldash.pdf}
      {\Huge$\meanasoc$} \quad \quad 
      \includegraphics[width=1.25in]{Figures/legendElements/dot.pdf}
      {\Huge$\meansoc$} \quad \quad
      \includegraphics[width=1.25in]{Figures/legendElements/ldashdot.pdf}
      {\Huge$\meanasoc = \meansoc$}
    \end{center}
\end{document}

### 2022-08-29 ms.tex
major figure updates

\begin{abstract} 
  Social learning is %essential to human survival---it 
  is a critical
  adaptation for dealing with uncertainty. However, uncertainty takes many forms.
  We identified four distinct and theoretically important types of uncertainty: temporal
  environmental variability, ambiguous payoffs, decision set size, and effective
  lifespans. We embed this framework in an evolutionary agent-based model to study conditions for the evolution of  social learning in populations with cognitively realistic individual learning
  capacities. % We used evolutionary agent-based modeling to track the evolution of social learning versus purely individual learning.
  %how often social, rather than individual,   learning evolved. 
  Agents could perform one of several behaviors, modeled as multi-armed bandit, to acquire payoffs.
  Individual learning used the softmax algorithm.  Social learning involved vertical and oblique transmission between generations, which served as a constraint on subsequent individual learning.  We found that different types of
  uncertainty had different effects. Temporal environmental variability suppressed
  social learning. Larger decision set sizes promoted social learning even
  when the environment changed frequently. Payoff
  ambiguity and lifespan had more complex effects on social learning, dependent on
  other uncertainty parameters.  This study shows the importance of clearly
  operationalizing uncertainty for understanding its role in social and cultural evolution.
% Social learning is essential to human survival. It is likely to evolve when it is more
% efficient than asocial, trial-and-error learning. Theoretically, social learning
% is adaptive for some uncertainty, but too much uncertainty makes
% social information unreliable. This fact is empirically
% supported across biology and human sciences. However, it is unclear how
% specific classes and types of uncertainty affect social
% learning evolution. Furthermore, existing models and experimental operationalizations of
% uncertainty are ambiguously related, and 
% tend to only consider a small number of behavioral
% choices.  Here we use evolutionary agent-based modeling to consolidate
% models of uncertainty in social learning evolution to address these issues.
% We model societies of agents who perform one of a number
% of possible behaviors to acquire payoffs in a time-varying environment.
% We identified complex patterns in social learning evolution depending on contextual
% uncertainty variables: environmental variability, ambiguous payoffs, 
% decision sets size, and effective lifespans. 
% Our work advances social learning evolution theory in a way that could help 
% guide human, non-human, and artificial intelligence towards optimal responses 
% to existential threats and new opportunities.
\footnote{This document contains
\total{citnum} references.}  
\end{abstract}

@@

\begin{figure}
  \caption{Social learning prevalence (y-axes) monotonically decreases as 
  environmental variabilty, $u$, increases (x-axes) in most uncertainty contexts. 
  Other uncertainty values $\pilow$ (rows), $B$ (columns), and $L$ (keys)
  shift and flatten the decrease from all-social-learner populations to all-asocial 
  populations.}
  \label{fig:mainResults}
  \centering
    \includegraphics[width=\textwidth]{{Figures/supplement/nagents=1000/mainResultsPlots.pdf}}
\end{figure}

@@

\begin{figure}
  \caption{Expected payoff rates for homogeneous populations of social or asocial learners. Dashed lines with no markers are expected payoff rates for a population of individual learners. Dashed lines with diamond markers are expected payoff rates to a population of social learners. Solid lines with circle markers are the actual average payoff rates for agents in our evolutionary simulations.  When $\pilow = 0.8$, all-social learner populations underperform
    since social learners are not subject to an initial evolutionary bottleneck
    due to the presence of asocial learners, and behavioral misinformation
    is allowed to circulate more freely than in the base case tested by 
    our model with an initially even distribution of social and asocial learners.
    \mt{Correct or nah?}
}
  \label{fig:payoffs}
  \centering
    \includegraphics[width=\textwidth]{Figures/supplement/fullMeanPrevNetPayoffs.pdf}
\end{figure}

@@

\begin{figure}
  \caption{Average number of generations ($\meanG$, y-axes) to fixation . 
    $\meanG$ increases and peaks around values of $u$ (x-axes) where
  $\meansoc \approx \meanasoc$. Note the y-axis ticks vary between plots,
indicating different overall time to fixation for different $(\pilow, B)$
pairs. Only extremal tested values of $L$ shown for clarity.} 
  \label{fig:steps}
\centering
    \includegraphics[width=\textwidth]{Figures/stepResultsPlots.pdf}
\end{figure}

@@

@@

\begin{figure}
  \ContinuedFloat
	\begin{subfigure}{\textwidth}
	  \caption{$N=100$}
	  \includegraphics[width=\textwidth]{Figures/mainResultsPlots.pdf}
	\end{subfigure}
\end{figure}

>>>>

\begin{abstract} 
  Social learning is a critical
  adaptation for dealing with uncertainty. However, uncertainty takes many forms.
  We identified four distinct and theoretically important types of uncertainty: temporal
  environmental variability, ambiguous payoffs, decision set size, and effective
  lifespans. We embed this framework in an evolutionary agent-based model to study conditions for the evolution of  social learning in populations with cognitively realistic individual learning
  capacities.  
  Agents could perform one of several behaviors, modeled as multi-armed bandit, to acquire payoffs.
  Individual learning used the softmax algorithm.  Social learning involved vertical and oblique transmission between generations, which served as a constraint on subsequent individual learning.  We found that different types of
  uncertainty had different effects. Temporal environmental variability suppressed
  social learning. Larger decision set sizes promoted social learning even
  when the environment changed frequently. Payoff
  ambiguity and lifespan had more complex effects on social learning, dependent on
  other uncertainty parameters.  This study shows the importance of clearly
  operationalizing uncertainty for understanding its role in social and cultural evolution.
% Social learning is essential to human survival. It is likely to evolve when it is more
% efficient than asocial, trial-and-error learning. Theoretically, social learning
% is adaptive for some uncertainty, but too much uncertainty makes
% social information unreliable. This fact is empirically
% supported across biology and human sciences. However, it is unclear how
% specific classes and types of uncertainty affect social
% learning evolution. Furthermore, existing models and experimental operationalizations of
% uncertainty are ambiguously related, and 
% tend to only consider a small number of behavioral
% choices.  Here we use evolutionary agent-based modeling to consolidate
% models of uncertainty in social learning evolution to address these issues.
% We model societies of agents who perform one of a number
% of possible behaviors to acquire payoffs in a time-varying environment.
% We identified complex patterns in social learning evolution depending on contextual
% uncertainty variables: environmental variability, ambiguous payoffs, 
% decision sets size, and effective lifespans. 
% Our work advances social learning evolution theory in a way that could help 
% guide human, non-human, and artificial intelligence towards optimal responses 
% to existential threats and new opportunities.
% \footnote{This document contains
% \total{citnum} references.}  
\end{abstract}

@@

\begin{figure}
  \caption{Social learning prevalence (y-axes) monotonically decreases as 
  environmental variabilty, $u$, increases (x-axes) in most uncertainty contexts. 
  Other uncertainty values $\pilow$ (rows), $B$ (columns), and $L$ (keys)
  shift and flatten the decrease from all-social-learner populations to all-asocial 
  populations.}
  \label{fig:mainResults}
  \centering
    \includegraphics[width=\textwidth]{Figures/mainResultsPlots.pdf}
\end{figure}

@@

@@

@@

\section{Expected payoffs}
\begin{figure}
  \caption{Expected payoff rates for homogeneous populations of social or asocial learners. Dashed lines with no markers are expected payoff rates for a population of individual learners. Dashed lines with diamond markers are expected payoff rates to a population of social learners. Solid lines with circle markers are the actual average payoff rates for agents in our evolutionary simulations.  When $\pilow = 0.8$, all-social learner populations underperform
    since social learners are not subject to an initial evolutionary bottleneck
    due to the presence of asocial learners, and behavioral misinformation
    is allowed to circulate more freely than in the base case tested by 
    our model with an initially even distribution of social and asocial learners.
    \mt{Correct or nah?}
}
  \label{fig:payoffs}
  \centering
    \includegraphics[width=\textwidth]{Figures/meanNetPayoffs.pdf}
\end{figure}

\section{Drift measured by expected generations to fixation}
\begin{figure}
  \caption{Average number of generations ($\meanG$, y-axes) to fixation . 
    $\meanG$ increases and peaks around values of $u$ (x-axes) where
  $\meansoc \approx \meanasoc$. Note the y-axis ticks vary between plots,
indicating different overall time to fixation for different $(\pilow, B)$
pairs. Only extremal tested values of $L$ shown for clarity.} 
  \label{fig:steps}
\centering
    \includegraphics[width=\textwidth]{Figures/stepResultsPlots.pdf}
\end{figure}

@@

\begin{figure}
  \ContinuedFloat
	\begin{subfigure}{\textwidth}
	  \caption{$N=100$}
	  \includegraphics[width=\textwidth]{Figures/supplement/nagents=100/mainResultsPlots.pdf}
	\end{subfigure}
\end{figure}
